---
title: "Analyze sampled wells"
output: github_document
---

```{r message=FALSE}
library(magrittr)
library(tidyverse)
```

Read data

```{r}

backend_file <- "output/subset_BR00106975/backend/subset_BR00106975.rds"

if (file.exists(backend_file)) {
  single_cell <- readRDS(backend_file)
  
} else {
  cytoplasm <- read_csv("output/subset_BR00106975/analysis/subset/Cytoplasm.csv")
  
  names(cytoplasm) <- str_c("Cytoplasm_", names(cytoplasm))
  
  nuclei <- read_csv("output/subset_BR00106975/analysis/subset/Nuclei.csv")
  
  names(nuclei) <- str_c("Nuclei_", names(nuclei))
  
  images <- 
    read_csv("output/subset_BR00106975/analysis/subset/Image.csv") %>%
    select(ImageNumber, Metadata_Plate, Metadata_Well, Metadata_Site)
  
  single_cell <- 
    nuclei %>% 
    inner_join(cytoplasm, 
               by = c("Nuclei_ObjectNumber" = "Cytoplasm_Parent_Nuclei",
                      "Nuclei_ImageNumber" = "Cytoplasm_ImageNumber")) %>%
    rename(ImageNumber = Nuclei_ImageNumber) %>%
    inner_join(images)
  
  metadata <- read_tsv("metadata/platemaps/CELLPAINTING.txt")
  
  names(metadata) <- str_c("Metadata_", names(metadata))
  
  single_cell %<>%
    inner_join(metadata, 
               by = c("Metadata_Well" = "Metadata_well_position")) %>%
    select(matches("Metadata"), everything())
  
  if (!dir.exists(dirname(backend_file))) {
      dir.create(dirname(backend_file), recursive = TRUE)
  }
  
  single_cell %>%
    saveRDS(backend_file)  
  
}

```

Report cell counts 

```{r cell_counts}
single_cell %>% 
  group_by(Metadata_Well, Metadata_plating_density, Metadata_condition) %>%
  tally() %>%
  ggplot(aes(as.factor(Metadata_plating_density), n)) + 
  geom_jitter(height = 0, width = .1) + 
  xlab("Plating density") +
  ylab("Cell count") +
  facet_wrap(~Metadata_condition)

single_cell %>% 
  group_by(Metadata_plating_density, Metadata_condition) %>%
  tally()
```
Plot distributions of cell area across all conditions

```{r nuclei_area_hist}
ggplot(single_cell,
       aes(Nuclei_AreaShape_Area)) +
  geom_histogram(binwidth = 30)

ggplot(single_cell %>% filter(Nuclei_AreaShape_Area > 2000),
       aes(Nuclei_AreaShape_Area)) +
  geom_histogram(binwidth = 30)
```
Filter out nuclei that seem like outliers based on area.
This is a bit of hack because in theory, very large nuclei may be informative.

```{r}
single_cell_filt <- 
  single_cell %>% 
  filter(Nuclei_AreaShape_Area < 2500)
```

How many cells are left? How many thrown out?

```{r}
single_cell_filt %>%
  nrow()

single_cell %>% 
  filter(Nuclei_AreaShape_Area >= 2500) %>%
  nrow()
```

```{r nuclei_area_violin}
single_cell_filt %>% 
  ggplot(aes(as.factor(Metadata_plating_density), Nuclei_AreaShape_Area)) + 
  geom_violin() +
  xlab("Plating density") +
  facet_wrap(~Metadata_condition)
```


To start with, select the lowest plating density (5000), and with glia for all further analysis. 
TODO:
- The `no_glia` condition looks odd - shouldn't the nuclei be larger?  

```{r}
single_cell_glia_5k <- 
  single_cell_filt %>% 
  filter(Metadata_plating_density == 5000 & 
           Metadata_condition == "glia")

```

Plot the distribution of nuclei size

```{r nuclei_area_hist_2}
ggplot(single_cell_glia_5k,
       aes(Nuclei_AreaShape_Area)) +
  geom_histogram(binwidth = 30)

ggplot(single_cell_glia_5k %>% filter(Nuclei_AreaShape_Area > 1000),
       aes(Nuclei_AreaShape_Area)) +
  geom_histogram(binwidth = 30)
```

Filter out nuclei that seem like outliers based on area.
This is a bit of hack because in theory, very large nuclei may be informative.

```{r}
single_cell_glia_5k_filt <- 
  single_cell_glia_5k %>% 
  filter(Nuclei_AreaShape_Area < 2000)
```

How many cells are left? How many thrown out?

```{r}
single_cell_glia_5k_filt %>%
  nrow()

single_cell_glia_5k %>% 
  filter(Nuclei_AreaShape_Area >= 2000) %>%
  nrow()
```

Plot the distribution of nuclei size of this filtered set

```{r nuclei_area_hist_filt_glia}
ggplot(single_cell_glia_5k_filt,
       aes(Nuclei_AreaShape_Area)) +
  geom_histogram(binwidth = 30)
```

Now fit two gaussians to the distribution of nucleus area.

The smaller nuclei are more likely to correspond to glia, while the larger ones to neurons

```{r}
library(mclust)

model <- Mclust(log(1+single_cell_glia_5k_filt$Nuclei_AreaShape_Area), G=2, model="V")

single_cell_glia_5k_filt %<>% mutate(Metadata_class = model$classification)

small_nuclei_index <- 
  single_cell_glia_5k_filt %>% 
  group_by(Metadata_class) %>% 
  summarise(mean_area = mean(Nuclei_AreaShape_Area)) %>%
  arrange(mean_area) %>% 
  slice(1) %>%
  extract2("Metadata_class") 

single_cell_glia_5k_filt %<>% 
  mutate(Metadata_class = 
           ifelse(Metadata_class == small_nuclei_index, "small nuclei", "large nuclei")
         )

```


```{r nuclei_area_hist_split}
ggplot(single_cell_glia_5k_filt,
       aes(x = Nuclei_AreaShape_Area, fill = Metadata_class)) +
  geom_histogram(binwidth = 20, position = "identity", alpha = 0.7)
```

Plot total DNA, using the same split (small nuclei, large nuclei) as above

```{r dna_hist_split} 
ggplot(single_cell_glia_5k_filt,
       aes(Nuclei_Intensity_IntegratedIntensity_DNA, fill = Metadata_class)) +
  geom_histogram(binwidth = 3, position = "identity", alpha = 0.7)
```

Plot total intensity in the ER channel (which overlaps with the GFP channel; GFP should be present only in neurons)

```{r er_hist_split}
ggplot(single_cell_glia_5k_filt,
       aes(Cytoplasm_Intensity_IntegratedIntensity_ER, fill = Metadata_class)) +
  geom_histogram(binwidth = 3, position = "identity", alpha = 0.7)
```
That has outliers, so filter

```{r er_hist_split_filt}
ggplot(single_cell_glia_5k_filt %>% filter(Cytoplasm_Intensity_IntegratedIntensity_ER < 350),
       aes(Cytoplasm_Intensity_IntegratedIntensity_ER, fill = Metadata_class)) +
  geom_histogram(binwidth = 2, position = "identity", alpha = 0.7)

```


```{r}
single_cell_glia_5k_filt_mat <- 
  single_cell_glia_5k_filt %>% 
  select(matches("^Nuclei_|^Cytoplasm_|Metadata_class"))

dim(single_cell_glia_5k_filt_mat)  
```

Next step: 

- Filter these 1432 features by removing redundant features

- Visualize the single cell data using UMAP

- Go back a few steps and decide what plating density to use, and whether comparing glia vs no glia is going to be meaningful in this dataset
