---
title: "Select images to print"
---

```{r message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(glue)
library(magrittr)
library(stringr)
library(tidyverse)
```

# Prepare load_data

```{r prepare_load_data, message=FALSE}
load_data <-
  list.files(
    "load_data_csv",
    pattern = "load_data.csv",
    full.names = T,
    recursive = T
  )  %>%
  map_df(function(filename) {
    df <- read_csv(filename, show_col_types = FALSE)
    df <- df %>% mutate(Metadata_Batch = str_split(filename, "/")[[1]][2])
  })

convert_to_s3_url <- function(pathname, filename)  {
  file.path(
      str_replace(pathname,
              "/home/ubuntu/bucket",
              "https://s3.amazonaws.com/imaging-platform"),
      filename
  )

}

channels <- c("DNA", "ER", "RNA", "AGP", "Mito")

for (channel in channels) {
  url_sym <- rlang::sym(str_c("URL_Orig", channel))

  path_sym <- rlang::sym(str_c("PathName_Orig", channel))

  file_sym <- rlang::sym(str_c("FileName_Orig", channel))

  load_data %<>%
    mutate(!!url_sym := convert_to_s3_url((!!path_sym), (!!file_sym)))

}

```

# Get metadata

```{r message=FALSE}

datasets <-
  tribble(
    ~batch, ~plate,
    "NCP_STEM_1", "BR_NCP_STEM_1",
    "NCP_PROGENITORS_1", "BR00127194",
    "2022_03_03_NCP_NEURONS_2_20x", "BR00132672",
    "2022_03_03_NCP_NEURONS_2_20x", "BR00132673",
    "2022-08-07_batch", "Plate1_PE_PP96",
    "2022-08-07_batch", "PE_PP_Plate2"
  )

metadata <-
  map2_df(datasets$batch,
          datasets$plate,
          function(batch, plate) {
            read_tsv(
              file.path("metadata", glue("{batch}/platemap/{plate}.txt")),
              show_col_types = FALSE
            ) %>%
              mutate(Batch = batch) %>%
              distinct()
          })

metadata %<>%
  rename(Plate = plate_map_name,
         Well = well_position)

names(metadata) <- str_c("Metadata_", names(metadata))
```

# Report plating densities

```{r}
metadata %>% group_by(Metadata_Batch, Metadata_Plate, Metadata_plating_density) %>% tally()
```


# Sample one well per cell line

```{r}
set.seed(5)

metadata_sampled <-
  metadata %>%
  group_by(Metadata_Batch, Metadata_line_ID, Metadata_line_condition, Metadata_plating_density) %>%
  sample_n(1) %>%
  ungroup()
```

# Select a fixed field (a.k.a site)

```{r}
fieldIDs <- c(3, 5)
```

# Select images to be printed

```{r}
filenames_header <- paste0("FileName_Orig", channels)

images <-
  load_data %>%
  select(Metadata_Batch, Metadata_Plate, Metadata_Well, Metadata_Row, Metadata_FieldID, matches("^URL_"), one_of(filenames_header)) %>%
  inner_join(metadata_sampled, by = c("Metadata_Batch", "Metadata_Plate", "Metadata_Well")) %>%
  filter(Metadata_FieldID %in% fieldIDs) %>%
  select(matches("^Metadata"), matches("^URL"), matches("^FileName_Orig"))

images_pivoted <-
  images %>%
  select(Metadata_Plate,  Metadata_Well, matches("^URL_")) %>%
  gather(Metadata_Channel, URL, -Metadata_Plate, -Metadata_Well) %>%
  mutate(filename = basename(URL)) %>%
  select(Metadata_Plate, Metadata_Well, Metadata_Channel, filename, URL)

images_pivoted %>%
  write_csv("data/sample_images.csv")

```

```{r}
images %>%
  select(matches("^Metadata"),  matches("^FileName_Orig")) %>%
  write_csv("data/sample_images_metadata.csv")
```

```{r}
images %>%
  distinct(Metadata_Batch, Metadata_Plate, Metadata_Well, Metadata_line_ID, Metadata_line_condition, Metadata_plating_density) %>%
  group_by(Metadata_Batch, Metadata_Plate, Metadata_line_ID, Metadata_line_condition, Metadata_plating_density) %>%
  tally() %>%
  group_by(n) %>%
  tally(name = "n_unique_conditions")
```

```{r}
metadata_sampled %>% count()
```


Run this on command line to download the images:

```{sh eval=FALSE}
IMAGE_DIR=/tmp/ncp

mkdir -p $IMAGE_DIR

cut -d"," -f1 data/sample_images.csv | grep -v Metadata_Plate| sort -u > /tmp/plates.txt

parallel -a /tmp/plates.txt --no-run-if-empty mkdir -p $IMAGE_DIR/{}

parallel \
 --header ".*\n" \
 -C "," \
 -a data/sample_images.csv \
 --eta \
 --joblog ${IMAGE_DIR}/download.log \
 wget -q -O ${IMAGE_DIR}/{1}/{4} {5}
```


# Generate outline URLs

Below, we create `URL_Outline` using `URL_OrigDNA` as the starting point.

Outlines are currently available only for `NCP_PILOT_3` and `NCP_PILOT_3B`

`URL_OrigDNA`:
https://s3.amazonaws.com/imaging-platform/projects/2019_05_28_Neuronal_Cell_Painting/workspace/analysis/NCP_PILOT_3/BR_NCP_PILOT_3/analysis/BR_NCP_PILOT_3-A18-3/r01c18f03p01-ch5sk1fk1fl1_Overlay.png

`URL_Outline`:
https://s3.amazonaws.com/imaging-platform/projects/2019_05_28_Neuronal_Cell_Painting/NCP_PILOT_3/images/BR_NCP_PILOT_3/Images/r01c18f03p01-ch5sk1fk1fl1.tiff

Note: Depending on how the Outline filename was defined the CellProfiler pipeline, you may need to choose a different channel as the starting point and also change the z plane.

E.g. if the pipeline was configured differently, we'd have different values in `?` for `r01c18f03p0?-ch?sk1fk1fl1_Overlay.png`

Only `NCP_PILOT_3B` and `NCP_PILOT_1` have 3 planes

```{r eval=FALSE}
load_data %>% distinct(Metadata_Batch, Metadata_PlaneID) %>% filter(Metadata_PlaneID > 1)
```

```{r eval=FALSE}
plane_batch <-
  tribble(~Metadata_Batch, ~ Metadata_PlaneID_outline,
           "NCP_PILOT_3B", 3,
           "NCP_PILOT_3", 1)

outline_images <-
  load_data %>%
  filter(Metadata_Batch %in% c("NCP_PILOT_3", "NCP_PILOT_3B")) %>%
  inner_join(plane_batch) %>%
  filter(Metadata_PlaneID_outline == Metadata_PlaneID) %>%
  select(
    Metadata_Batch,
    Metadata_Plate,
    Metadata_Well,
    Metadata_Row,
    Metadata_FieldID,
    matches("^URL_")
  ) %>%
  inner_join(metadata_sampled,
             by = c("Metadata_Batch", "Metadata_Plate", "Metadata_Well")) %>%
  filter(Metadata_FieldID %in% fieldIDs) %>%
  select(matches("^Metadata"), matches("^URL")) %>%
  mutate(
    URL_Outline =
      file.path(
        dirname(dirname(dirname(dirname(
          dirname(URL_OrigDNA)
        )))),
        "workspace",
        "analysis",
        Metadata_Batch,
        Metadata_Plate,
        "analysis",
        str_c(Metadata_Plate, "-", Metadata_Well, "-", Metadata_FieldID),
        str_replace(basename(URL_OrigDNA), "\\.tiff", "_Overlay.png")
      )
  ) %>%
  select(matches("^Metadata"), URL_Outline)

outline_images %>%
  write_csv("data/sample_images_outline.csv")
```

```{r eval=FALSE, results='asis'}
outline_images %>%
  mutate(URL_Outline = str_c("[Outline](", URL_Outline, ")")) %>%
  select(Metadata_Batch, Metadata_plating_density, Metadata_line_condition, Metadata_compound_ID, URL_Outline, everything()) %>%
  knitr::kable()
```
