---
title: "Inspect profiles"
author: "Shantanu Singh"
date: "June 2021"
---

# Load libraries

```{r message=FALSE}
library(ggplot2)
library(glue)
library(magrittr)
library(tidyverse)
```

# Read profiles

```{r message=FALSE}
batch_id <- "NCP_PILOT_3B"

platemap <- "MAtt_ICC_test"

plate_id <- "MAtt_ICC_test"

profiles_map2 <- 
  read_csv(file.path("profiles", batch_id, plate_id, paste0(plate_id, "_augmented.csv.gz")), 
           progress = FALSE)
```

```{r message=FALSE}
batch_id <- "NCP_PILOT_3"

platemap <- "BR_NCP_PILOT_3"

plate_id <- "BR_NCP_PILOT_3"

profiles_cp <- 
  read_csv(file.path("profiles", batch_id, plate_id, paste0(plate_id, "_augmented.csv.gz")), 
           progress = FALSE)
```


```{r}
shorten_skel_feature_name <- function(s) 
  s %>% 
  str_remove("Nuclei_ObjectSkeleton_") %>% 
  str_remove("_CellBodiesPlusNeuritesSkel")

get_skel_features <- function(profiles) 
  profiles %>% 
  select(matches("Metadata_|Skel$")) %>% 
  select(-matches("_mito_skel$")) %>%
  rename_with(shorten_skel_feature_name)
  

count_wells <- function(profiles) 
  profiles %>% group_by(across(matches("plating_density|line_condition|compound_ID"))) %>% tally()
```

# Report statistics

## Global

### Counts 

Count wells per unique perturbations

```{r rows.print=40}
count_wells(skel_map2)

count_wells(skel_cp)
```

### Branching metrics

Get branching metrics

From http://cellprofiler-manual.s3.amazonaws.com/CellProfiler-3.0.0/modules/measurement.html#measureobjectskeleton

- `NumberTrunks`: The number of trunks. Trunks are branchpoints that lie within the seed objects
- `NumberNonTrunkBranches`: The number of non-trunk branches. Branches are the branchpoints that lie outside the seed objects.
- `NumberBranchEnds`: The number of branch end-points, i.e, termini.
- `TotalObjectSkeletonLength`: The length of all skeleton segments per object.


```{r}
skel_map2 <- get_skel_features(profiles_map2)

skel_cp <- get_skel_features(profiles_cp)
```

Plot these features

```{r}
skel_map2 %>%
  pivot_longer(-matches("Metadata")) %>%
  ggplot(aes(value)) + geom_histogram(bins = 50) + facet_wrap(~name, scales = "free")
```


```{r}
skel_cp %>%
  pivot_longer(-matches("Metadata")) %>%
  ggplot(aes(value)) + geom_histogram(bins = 50) + facet_wrap(~name, scales = "free")
```

```{r}
variables <- names(skel_cp) %>% str_subset("Metadata_", negate = TRUE)
skel_cp %>% 
  cytominer::mark_outlier_rows(variables, skel_cp) %>% 
  filter(!is_outlier) %>%
  select(-is_outlier) %>%
  pivot_longer(-matches("Metadata")) %>%
  ggplot(aes(value)) + geom_histogram(bins = 50) + facet_wrap(~name, scales = "free")
  
```
```{r}
skel_cp <- 
  skel_cp %>% 
  cytominer::mark_outlier_rows(variables, skel_cp) %>% 
  filter(!is_outlier) %>%
  select(-is_outlier) 

skel_cp %>%
  count()
```

## Per condition

### Branching metrics

```{r fig.height=7, fig.width=5}
skel_map2 %>%
  pivot_longer(-matches("Metadata")) %>%
  ggplot(aes(Metadata_line_condition, value)) + geom_boxplot() + 
  facet_grid(name~Metadata_plating_density, scales = "free")
```

```{r fig.height=10, fig.width=7}
skel_cp %>%
  pivot_longer(-matches("Metadata")) %>%
  ggplot(aes(Metadata_compound_ID, value, fill=Metadata_line_condition)) + 
  geom_boxplot() + 
  facet_grid(name~Metadata_plating_density, scales = "free") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(legend.position="bottom")
```


```{r}
skel_cp <- 
  skel_cp %>% 
  select(
    any_of(c("Metadata_plating_density", "Metadata_line_condition", "Metadata_compound_ID")), 
    all_of(variables)) %>%
  mutate(Metadata_compound_ID = replace_na(Metadata_compound_ID, "Untreated_CP"))

skel_map2 <- 
  skel_map2 %>% 
  select(
    any_of(c("Metadata_plating_density", "Metadata_line_condition", "Metadata_compound_ID")), 
    all_of(variables)) %>%
  mutate(Metadata_compound_ID = "Untreated_MAP2") %>%
  select(matches("Metadata"), everything())
```

```{r}
skel <-
  bind_rows(skel_cp, skel_map2)
```

```{r fig.height=10, fig.width=7}
skel %>%
  pivot_longer(-matches("Metadata")) %>%
  ggplot(aes(Metadata_compound_ID, value, fill=Metadata_line_condition)) + 
  geom_boxplot() + 
  facet_grid(name~Metadata_plating_density, scales = "free") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(legend.position="bottom") +
  ggtitle("Branching metrics from Cell Painting and MAP2 images", 
          subtitle = "All compounds are from Cell Painting images")
```
```{r fig.height=10, fig.width=7}
skel %>%
  pivot_longer(-matches("Metadata")) %>%
  filter(str_detect(Metadata_compound_ID, "Untreated")) %>%
  ggplot(aes(Metadata_compound_ID, value, fill=Metadata_line_condition)) + 
  geom_boxplot() + 
  facet_grid(name~Metadata_plating_density, scales = "free") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(legend.position="bottom") +
  ggtitle("Branching metrics from Cell Painting and MAP2 images", 
          subtitle = "Only untreated wells")
```